New lower bounds for Schur and weak Schur numbers

### 2021-10-07 article/article.tex
modifs sans WS

We exhibit new templates for Schur and Ramsey numbers. We also generalize the concept of template to weak Schur numbers. 
These templates provide new general inequalities of the form \(S(n+k) \geqslant a S(n) + b\) and \(\WS(n+k) \geqslant a S(n) + b\) 
as well as new lower bounds for both Schur and weak Schur numbers. We also give a new lower bound \(\WS(6) \geqslant 646\) with 
a method that can yield slight improvements for fixed values of \(n\) over lower bounds obtained with the inequality 
\(\WS(n+1) \geqslant 4 S(n) + 2\).

Given that the introduction of templates is quite recent, we expect the bounds to be improved as this special type of partition 
becomes better understood and larger templates are found. One may try to find better templates by using heuristics or 
approximation algorithms, such as Monte-Carlo algorithms for instance.

The best lower bound \(\WS(6) \geqslant 583\) achieved with a computer search using Monte-Carlo methods 
\cite{EliahouBook} is significantly lower than those obtained with a template-like structure (\(\WS(6) \geqslant 642\) \cite{RowleyWS}, 
\(\WS(6) \geqslant 646\) in the present article). We have evidence that 583 is the maximal value in the search space considered by 
\cite{EliahouBook,Bouzy2015AnAP,Rafilipojaona}. Therefore, it questions the assumption that large partitions for \(\WS(n+1)\) 
start with large partitions for \(\WS(n)\). It also indicates that \(\WS(5)\) might need further investigation since the current lower 
bound \(\WS(5) \geqslant 196\) \cite{ELIAHOU2012175} was obtained by considering the same type of search space.

>>>>

We have produced new templates for Schur and Ramsey numbers. We also have generalized the concept of template to weak Schur numbers. 
These templates provide new general inequalities of the form \(S(n+k) \geqslant a S(n) + b\) and \(\WS(n+k) \geqslant a S(n) + b\) 
as well as new lower bounds for both Schur and weak Schur numbers. We have given a new lower bound \(\WS(6) \geqslant 646\) with 
a method that can yield slight improvements for fixed values of \(n\) over lower bounds obtained with the inequality 
\(\WS(n+1) \geqslant 4 S(n) + 2\).

Given that the introduction of templates is quite recent, we expect the bounds to be improved as this special type of partition 
becomes better understood and larger templates are found. One may try to produce better templates by using heuristics or 
approximation algorithms, such as Monte-Carlo algorithms for instance.

The best lower bound \(\WS(6) \geqslant 583\) achieved with a computer search using Monte-Carlo methods 
\cite{EliahouBook} is significantly lower than those obtained with a template-like structure: \(\WS(6) \geqslant 642\) \cite{RowleyWS}, 
\(\WS(6) \geqslant 646\). We have evidence that 583 is the maximal value in the search space considered by 
\cite{EliahouBook,Bouzy2015AnAP,Rafilipojaona}. Therefore, it questions the assumption that large partitions for \(\WS(n+1)\) 
start with large partitions for \(\WS(n)\). It also indicates that \(\WS(5)\) might need further investigation since the current lower 
bound \(\WS(5) \geqslant 196\) \cite{ELIAHOU2012175} was obtained by considering the same type of search space.

### 2021-10-07 article/introduction.tex
modifs sans WS

We are interested in partioning the set of integers \(\{1, ..., p\}\) into \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we add the hypothesis \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there exists a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Values of \(S(n)\) and \(\WS(n)\) are known for small \(n\) only.

@@

Up to recently, the most efficient generic construction for Schur numbers was given by Abbott and Hanson 
\cite{AbbottHanson} in 1972 with a recursive construction. It gave the best lower bounds for all sufficiently large 
numbers. No equivalent was known for weak Schur numbers and as a result the best known partitions for large 
weak Schur numbers did not use the weakly sum-free hypothesis.

\par
As for smaller weak Schur numbers, the best lower bounds were obtained by conducting a computer search. Eliahou 
\cite{EliahouBook}, Bouzy \cite{Bouzy2015AnAP} and Rafilipojaona \cite{Rafilipojaona} improved the lower 
bounds with Monte-Carlo methods. This was the main approach during the past decade.

\par
In 2020, Rowley introduced the notion of templates for Schur and Ramsey numbers \cite{RowleyRamsey} which 
generalizes Abbott and Hanson's construction and produces new lower bounds (and inequalties) for Schur numbers. 
He also provided two inequalities for weak Schur numbers \cite{RowleyWS} that yield significant improvements 
over previous lower bounds. Besides, these two inequalities do use the \textit{weakly} sum-free hypothesis.

\par The hereunder tables recap the lower bounds respectively for Schur and weak Schur numbers. Gray cells 
indicate the exact values.

@@

Our main contribution is a generalization of the concept of template to weak Schur numbers. Our templates 
provide new lower bounds (and inequalities) for weak Schur numbers. This construction also includes as a special case 
a construction similar to Abbott and Hanson's \cite{AbbottHanson}, but this time for \textit{weak} Schur numbers.

\par
In Section \ref{Schur}, we explain Rowley's template-based construction in the context of 
Schur numbers. Then, we give new templates, thus providing new lowers bounds and inequalities as well as 
showing that the growth rates for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
In Section \ref{WeakSchur}, we generalize the concept of template to weak Schur numbers 
and provide new lower bounds for weak Schur numbers. We then use a different approach and give a new 
lower bound for \(\WS(6)\).

\par
We now introduce notations and definitions we use throughout this article.

@@

We start by defining sum-free and weakly sum-free subsets to introduce regular and weak Schur numbers. 
We denote \(\mathbb{N} = \{0, 1, 2, ...\}\) and \(\mathbb{N}^* =\mathbb{N} \backslash \{0\}\).

@@

\par
Schur proved in \cite{Schur1917} that given a number of subsets \(n\), there exists a value of \(p\)
such that there exists no partition of \([\![1,q]\!]\) into \(n\) sum-free subsets for any \(q \geqslant p\). A similar
property holds for weakly sum-free subsets \cite{Irving1973}. These observations lead to the following definitions.

\begin{definition}
Let \(n \in \mathbb{N}^*\). There exists a largest integer denoted by \(S(n)\) such that \([\![1, S(n)]\!]\) can be
 partitioned into \(n\) sum-free subsets. \(S(n)\) is called the \textit{\(n\)\textsuperscript{th} Schur number}.
\end{definition}

\begin{definition}
Let \(n \in \mathbb{N}^*\). There exists a largest integer denoted by \(\WS (n)\) such that \([\![1, \WS (n)]\!]\) 
can be partitioned into \(n\) weakly sum-free subsets. \(\WS(n)\) is called the \(n\)\textsuperscript{th} weak Schur 
number.
\end{definition}

Given a partition of \([\![1, p]\!]\) into \(n\) subsets, we generally denote these subsets \(A_1, ..., A_n\). We also denote
\(m_i = \min(A_i)\). By ordering the subsets, we mean assuming that \(m_1 < ... < m_n\). However, if not specified we do 
not make this hypothesis since we do not always consider partitions in which every subset plays a symmetric role.

\begin{definition}
We sometimes refer to a partitition as a coloring. The coloring associated to a partition \(A_1, ..., A_n\) of 
\([\![1, p]\!]\) is the function \(f\) such that \(\forall x \in [\![1, p]\!], x \in A_{f(x)}\). Likewise, the partition associated to
a coloring \(f\) of \([\![1, p]\!]\) with \(n\) colors is \(\forall c \in [\![1, n]\!], A_c = f^{-1}(c)\).
\end{definition}

>>>>

We are interested in partioning the set of integers \(\{1, ..., p\}\) into \(n\) subsets such that there is no 
subset containing three integers \(x\), \(y\) and \(z\) verifying \(x + y = z\). We say these subsets are 
\textit{sum-free}. If we lessen the above constraint by only considering integers \(x,y\) such that \(x \neq y\), we say the subsets are \textit{weakly sum-free}. The 
greatest \(p\) for which there is a partition into \(n\) sum-free subsets is called the \(n^{\text{th}}\) Schur 
number and is denoted \(S(n)\) \cite{Schur1917}. Likewise for weakly sum-free partitions we define \(\WS(n)\) 
the \(n^{\text{th}}\) weak Schur number \cite{Irving1973}. Values of \(S(n)\) and \(\WS(n)\) are known for small \(n\) only.

@@

Up to recently, the most efficient generic construction for Schur numbers was given by Abbott and Hanson 
\cite{AbbottHanson} in 1972 with a recursive construction. It gave the best lower bounds for all sufficiently large 
numbers. No equivalent was known for weak Schur numbers and, as a result, the best known partitions for large 
weak Schur numbers did not use the \textit{weakly} sum-free property.

\par
As for smaller weak Schur numbers, the best lower bounds were obtained by computations. Eliahou 
\cite{EliahouBook}, Bouzy \cite{Bouzy2015AnAP} and Rafilipojaona \cite{Rafilipojaona} improved the lower 
bounds with Monte-Carlo methods. This was the main approach during the past decade.

\par
In 2020, Rowley introduced the notion of templates for Schur and Ramsey numbers \cite{RowleyRamsey} which 
generalizes Abbott and Hanson's construction and produces new lower bounds (and inequalties) for Schur numbers. 
He also provided two inequalities for weak Schur numbers \cite{RowleyWS}. Besides, these two inequalities do use the \textit{weakly} sum-free property.

\par The hereunder tables recap the lower bounds for Schur and weak Schur numbers, respectively. Gray cells 
indicate the exact values.

@@

Our main contribution is a generalization of the concept of template to weak Schur numbers. Our templates 
provide new lower bounds (and inequalities) for weak Schur numbers. As a special case, our construction also includes  
a construction similar to Abbott and Hanson's \cite{AbbottHanson}, but for \textit{weak} Schur numbers this time.

\par
In Section \ref{Schur}, we explain Rowley's template-based construction for 
Schur numbers. Then, we give new templates, thus providing new lowers bounds and inequalities as well as 
showing that the growth rates for both Schur and Ramsey numbers \(R_n(3)\) exceed 3.28.

\par
In Section \ref{WeakSchur}, we generalize the concept of template to weak Schur numbers 
and provide new lower bounds for weak Schur numbers. Then, we use a different approach and give a new 
lower bound for \(\WS(6)\).

\par
Now, we introduce notations and definitions used throughout this article.

@@

We start by defining sum-free and weakly sum-free subsets to introduce regular and weak Schur numbers. 
The set of positive natural numbers is denoted by \(\mathbb{N}^* =\mathbb{N} \backslash \{0\}\).

@@

\par
Schur proved \cite{Schur1917} that given a number of subsets \(n\), there is a value of \(p\)
such that \([\![1,q]\!]\) cannot be partionned into \(n\) sum-free subsets for \(q \geqslant p\). A similar
property holds for weakly sum-free subsets \cite{Irving1973}. These observations lead to the following definitions
written for \(n \in \mathbb{N}^*\):

\begin{definition}
There is a largest integer denoted by \(S(n)\) such that \([\![1, S(n)]\!]\) can be
 partitioned into \(n\) sum-free subsets. \(S(n)\) is called the \textit{\(n\)\textsuperscript{th} Schur number}.
\end{definition}

\begin{definition}
There is a largest integer denoted by \(\WS (n)\) such that \([\![1, \WS (n)]\!]\) 
can be partitioned into \(n\) weakly sum-free subsets. \(\WS(n)\) is called the \(n\)\textsuperscript{th} weak Schur 
number.
\end{definition}

Subsets of an \(n\)-partition are denoted \(A_1, ..., A_n\). The smallest element of \(A_i\) is denoted  
\(m_i = \min(A_i)\). By ordering the subsets, we mean assuming that \(m_1 < ... < m_n\). However, if not specified we do 
not make this assumption since we do not always consider partitions in which every subset plays a symmetric role.

A partition may be seen as a number coloring.

\begin{definition}
The coloring associated to a partition \(A_1, ..., A_n\) of 
\([\![1, p]\!]\) is the function \(f\) such that \(\forall x \in [\![1, p]\!], x \in A_{f(x)}\). Likewise, the partition associated to
a coloring \(f\) of \([\![1, p]\!]\) with \(n\) colors is \(\forall c \in [\![1, n]\!], A_c = f^{-1}(c)\).
\end{definition}

### 2021-10-07 article/schur.tex
rajout d'une lettre
modifs sans WS

In this section, we use Rowley's template-based constructions \cite{RowleyRamsey} in the context of Schur 
numbers. In order to improve lower bounds for Schur and Ramsey numbers, he introduces special sum-free 
partitions verifying some additional properties which can be extended using a method generalizing Abbott and 
Hanson's construction \cite{AbbottHanson}. He named these partitions "templates", and we keep this name in 
the entire article. We then find new templates and use them to provide new lower bounds for Schur numbers.

@@

We begin by introducing \textit{S-templates}, standing for \textit{"Schur templates"}. The idea is to consider the first line
of \hyperref[figure:1]{Figure 1} not as a combination of two blocs anymore but as a whole, single construction. A S-template 
is then defined as a new object filling the role of the first line but with less, yet sufficient, constraints which allow for an 
expansion of the partition using a sum-free partition.

\begin{definition}
Let \((p,n) \in (\mathbb{N}^*)^2\). A S-template with width \(p\) and \(n\) colors is defined as a partition of 
\([\![1,p]\!]\) into \(n\) sum-free subsets \(A_1, A_2, ..., A_n\) verifying for all subsets but \(A_n\)
\[
\forall i \in [\![1, n-1]\!], \forall (x,y) \in A_i^2, x+y > p
\Longrightarrow x+y-p \notin A_i.
\]
\end{definition}

Here \(n\) can be seen as a "special" color in the sense that we do not impose this above additionnal constraint on this color. 
However, note that \(n\) is not necessarily the last color by order of appearance, any color can play this role.

@@

\begin{proof}[\textsc{Proof.}]
The upper bound comes from the fact that a S-template with width \(p\) and \(n\) colors is also a partition of \([\![1, p]\!]\) 
into \(n\) sum-free subsets. The lower bound comes from Abbott and Hanson's construction \cite{AbbottHanson}. \\
\end{proof}

@@

We start by reminding the explicit construction of a sum-free partition with the use of a S-template and another sum free partition. 
We rephrase in terms of Schur numbers this contruction stated by Rowley in the context of Ramsey numbers \cite{RowleyRamsey}.

\begin{theorem}
\label{thm:Stemp}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a \(S\)-template with width \(q\) and \(n+1\) colors 
	and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets then there exists a partition of \([\![1,pq+m_{n+1}
	-1]\!]\) into \(n+k\) sum-free subsets. \(m_{n+1}\) denotes the minimum element colored with the special color in the \(S\)-template.
\end{theorem}

Setting \(q = S^+(n+1)\) and \(p = S(k)\) yields the following corollary.

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. They vertically 
extend a sum-free partition by repeating it and they use an other sum-free partition to color the other half according 
to the line number. This way the "blocks" act as safe areas for each other. We give here an example for \(p = 4\), 
\(q = 9\), \(n = 2\) and \(k = 2\) showing that \(S(2 + 2) \geqslant S(2) \left(2 S(2) + 1\right) + S(2)\), both with Abbott and 
Hanson's construction and with a S-template which is not included in Abbott and Hanson's construction. In both cases, 
the special color is grey.

@@

\setlength{\tabcolsep}{6pt}
\caption{Visualisation of an Abbott and Hanson construction}
\end{center}
\end{figure}
In the general construction with S-templates, the special color no longer necessarily contains consecutive 
numbers. However, the special color is still replaced by the colors of the sum-free partition according to the 
line number and the other colors are still vertically extended.

@@

We now proceed to prove Theorem \ref{thm:Stemp}.

\begin{proof}[\textsc{Proof.}]
	Denote by \(f\) the coloring associated to the \(S\)-template with width \(q\) and \(g\) the one associated 
	to the sum-free partition of \([\![1,p]\!]\); where \(f : [\![1,q]\!] \longrightarrow [\![1,n+1]\!]\) and 
	\(g : [\![1,p]\!] \longrightarrow [\![1,k]\!]\).

@@

	For \(x \in [\![1,pq+m_{n+1}-1]\!]\), we write \(x = (\alpha - 1) q + u\) for certain integers \(\alpha \in \mathbb{N}^*\) and 
	\(u \in [\![1,q]\!]\). This decomposition is of course unique. \(\alpha\)  can be interpreted as the row number of \(x\) and 
	\(u\) as the collumn number of \(x\). A new coloring is define as follows:
	\[
	\begin{array}{c c c l}
		h : & [\![1,pq+m_{n+1}-1]\!] & \longrightarrow & [\![1,n+k]\!], \\
		& x & \longmapsto & 
		\left\{ \begin{array}{l l}
			f(u), & \text{if}~f(u) \leqslant n, \\
			n + g(\alpha), & \text{if}~f(u) = n + 1. \\
		\end{array} \right.
	\end{array}
	\]

@@

	\par
	We now prove that \(h\) is a sum-free coloring. Let \(x,y \in [\![1,pq + m_{n+1}-1]\!]\) such that \(h(x) = h(y)\) 
	and \(x+y \leqslant pq+m_{n+1}-1\). We claim that \(h(x+y) \neq h(x)\). We write \(x = (\alpha - 1) q + u\) and 
	\(y = (\beta - 1) q + v\) where \(\alpha, \beta \in \mathbb{Z}\) and \(u, v \in [\![1,q]\!]\). Two cases are to be
	distinguished according to the value of \(h(x)\). \\

@@

	\noindent \underline{\textbf{Case 2:} \(h(x) \geqslant n + 1\)}
	\par
	Now we have \(h(x) = n + g(\alpha) = 
	n + g(\beta) = h(y)\), hence \(g(\alpha) = g(\beta)\). As in the first case, distinguish between two cases according 
	to the value of \(x + y\).

@@

@@

Then, for every \(n \in \mathbb{N}^*\), by using on the last row the coloring \(x \longmapsto g(x - q S(n))\), we have
\[ S(n+k) \geqslant q S(k) + b.\]
\end{proposition}
This proposition corresponds to the fact that the hypotheses made on the coloring of the last row can be weakened because 
the constraints in the definition of a S-template prevent a number from both being the sum of two of numbers of its own color 
as well as creating a sum in its own color, the later of which is not always relevant on the last row given that it is not completely filled.
As a result, we can change the coloring on \([\![ q S(k) + 1,  q S(k) + b]\!]\) (the last row) to extend the previous partitions.
\newline

@@

A S-template with \(n\) colors is said to be symmetric if the partition into \(n\) sum-free subsets derived 
from this template by applying the extension procedure with a sum-free partition of length 1 is symmetric. 
\end{definition}

We produced S-templates using a SAT solver, hence providing lower bound on \(S^+\) and inequalities 
of the type \(S(n+k) \geqslant a S(n) + b\). We sought templates providing the largest value of 
\((a, b)\) (in the lexicographic order). When the number of colors exceeded five, in order to reduce the search space we 
only looked for symmetric S-templates, we assumed that the special color was the last color to appear and we constrained 
the \(m_c\)'s out of being too small. Details concerning the encoding as a SAT problem can be found in \cite{Heule2017}.

\par
The following six inequalities are given by the current best S-templates with \(n \leqslant 7\) colors.

\begin{align}
	S(n + 1) \geqslant &~3\,S(n)  + 1 \label{S(n+1)}\\
	S(n + 2) \geqslant &~9\,S(n)  + 4 \label{S(n+2)}\\
	S(n + 3) \geqslant &~33\,S(n) + 6 \label{S(n+3)}
\end{align}

@@

\par The \hyperref[S-templates]{templates} corresponding to inequalities (\ref{S(n+3)}), (\ref{S(n+4)}) and 
(\ref{S(n+5)}) can be found in the appendix.

\par
Inequalities (\ref{S(n+1)}), (\ref{S(n+2)}) and (\ref{S(n+3)}) cannot be further improved (with this definition of S-template). 
Inequality (\ref{S(n+4)}) cannot be further improved by only searching for symmetric S-templates whose special color is the 
last in the order of apparition (and with a multiplicative factor less than or equal to 118). Inequality (\ref{S(n+5)}) can most 
likely be further improved but the improvement probably will not be substantial. Finally, a S-template corresponding to inequality 
(\ref{S(n+6)}) was found by extending into a S-template the Schur 6 partition used in \cite{rowley2021improved} (owing to the 
size of this template, its interest is limited and since it can easily be derived from above mentioned partition, it is not given in 
the appendix). Although we could not find a better S-template with seven colors, inequality (\ref{S(n+6)}) can definitely be 
improved by a wide margin.

@@

Except for \(S(8)\), \(S(9)\) and \(S(13)\), the best lower bounds are obtained thanks to
the fifth inequality \( S(n+5) \geqslant 380 \, S(n) + 148\). The table
doesn't go any further, but the same inequality allows to improve the
lower bounds for every \( n \geqslant 15 \).

@@

In this section, we first formalized Rowley's template-based constructions \cite{RowleyRamsey} in the context of Schur numbers 
by introducing S-templates as well as a new sequence, \(S^+\). We provided relations between \(S^+\) and \(S\) then 
stated Rowley's construction method in the context of Schur numbers. We found new S-templates allowing us to obtain 
new lower bounds for schur numbers. One may notice that we mostly gave only lower bounds for \(S^+\). It should be possible to 
find better S-templates by making different assumptions or using a different method (Monte-Carlo methods for instance).

\par
In the next section, we provide similar results for weak Schur numbers. We introduce WS-templates and a corresponding sequence, 
\(\WS^+\). We then derive similar relations and a construction method allowing us to find new lower bounds for weak Schur numbers.

>>>>

We use Rowley's template-based constructions \cite{RowleyRamsey} for Schur 
numbers. To improve lower bounds for Schur and Ramsey numbers, he has introduced special sum-free 
partitions verifying some additional properties which can be extended using a method generalizing Abbott and 
Hanson's construction \cite{AbbottHanson}. He called these partitions "templates", and we keep this nomination. 
Then, we propose new templates and use them to produce new lower bounds for Schur numbers.

@@

We begin by introducing \textit{S-templates}, standing for \textit{"Schur templates"}. The idea is to consider the first line
of \hyperref[figure:1]{Figure 1} not as a combination of two blocs, like in \cite{AbbottHanson} but as a whole, single construction. 
An S-template is then defined as a new object playing the role of the first line but with less, yet sufficient, 
constraints which allow for an expansion of the partition using a sum-free partition.

\begin{definition}
Let \((p,n) \in (\mathbb{N}^*)^2\). An S-template with \(n\) colors is defined as a partition of 
\([\![1,p]\!]\) into \(n\) sum-free subsets \(A_1, A_2, ..., A_n\) verifying for all subsets but \(A_n\)
\begin{equation}
\label{eq:constemp}
\forall (x,y) \in A_i^2, x+y > p
\Longrightarrow x+y-p \notin A_i.
\end{equation}
The integer \(p\) is called the \textit{width} of the template.
\end{definition}

Here \(n\) can be seen as a "special" color in the sense that it is no constrained by \ref{eq:constemp}.
However, note that \(n\) is not necessarily the last color by order of appearance, any color can play this role.

@@

\begin{proof}[\textsc{Proof.}]
The upper bound comes from the fact that an S-template with width \(p\) and \(n\) colors is also a partition of \([\![1, p]\!]\) 
into \(n\) sum-free subsets. The lower bound comes from Abbott and Hanson's construction \cite{AbbottHanson}. \\
\end{proof}

@@

We start by reminding the explicit construction of a sum-free partition with the use of an S-template and another sum free partition. 
We rephrase in terms of Schur numbers this contruction stated by Rowley in the context of Ramsey numbers \cite{RowleyRamsey}.

\begin{theorem}
\label{thm:Stemp}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there is an S-template with width \(q\) and \(n+1\) colors 
	and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets then there is a partition of \([\![1,pq+m_{n+1}
	-1]\!]\) into \(n+k\) sum-free subsets. \(m_{n+1}\) denotes the minimum element colored with the special color in the \(S\)-template.
\end{theorem}

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. They 
extended a sum-free partition vertically by repeating it. Next, they used another sum-free partition to color the other half according 
to the line number, as in \hyperref[figure:1]{Figure 1}. We give an example for \(p = 4\), 
\(q = 9\), \(n = 2\) and \(k = 2\) showing that \(S(2 + 2) \geqslant S(2) \left(2 S(2) + 1\right) + S(2)\), both with Abbott and 
Hanson's construction and with an S-template which is not included in Abbott and Hanson's construction. In both cases, 
the special color is grey.

@@

\setlength{\tabcolsep}{6pt}
\caption{Visualisation of an Abbott and Hanson construction}
\end{center}
\end{figure}
In the general construction with S-templates, the special color may not paint consecutive 
numbers any longer. However, the special color is still replaced by the colors of the sum-free partition according to the 
line number and the other colors are still vertically extended.

@@

Now, we prove Theorem \ref{thm:Stemp}.

\begin{proof}[\textsc{Proof.}]
	Let \(f\) and \(g\) be colorings associated to the S-template with width \(q\) and the sum-free partition of 
	\([\![1,p]\!]\), respectively: \(f : [\![1,q]\!] \longrightarrow [\![1,n+1]\!]\) and 
	\(g : [\![1,p]\!] \longrightarrow [\![1,k]\!]\).

@@

	For \(x \in [\![1,pq+m_{n+1}-1]\!]\), we write \(x = (\alpha - 1) q + u\) for certain integers \(\alpha \in \mathbb{N}^*\) and 
	\(u \in [\![1,q]\!]\). Obviously, this decomposition unique. Coefficients \(\alpha\) and \(u\) may be interpreted as the row index of \(x\) and 
	as line and column coordinates of \(x\) in the template. A new coloring is defined as:
	\[
	\begin{array}{c c c l}
		h : & [\![1,pq+m_{n+1}-1]\!] & \longrightarrow & [\![1,n+k]\!], \\
		& x & \longmapsto & 
		\left\{ \begin{array}{l l}
			f(u), & \text{if}~f(u) \leqslant n, \\
			n + g(\alpha), & \text{if}~f(u) = n + 1. \\
		\end{array} \right.
	\end{array}
	\]

@@

	\par
	Now, we prove that \(h\) is a sum-free coloring. Let \(x,y \in [\![1,pq + m_{n+1}-1]\!]\) such that \(h(x) = h(y)\) 
	and \(x+y \leqslant pq+m_{n+1}-1\). We claim that \(h(x+y) \neq h(x)\). We write \(x = (\alpha - 1) q + u\) and 
	\(y = (\beta - 1) q + v\) where \(\alpha, \beta \in \mathbb{Z}\) and \(u, v \in [\![1,q]\!]\). Two cases are to be
	distinguished according to the value of \(h(x)\). \\

@@

	\noindent \underline{\textbf{Case 2:} \(h(x) \geqslant n + 1\)}
	\par
	Now we have \(h(x) = n + g(\alpha) = 
	n + g(\beta) = h(y)\), hence \(g(\alpha) = g(\beta)\). As above, we distinguish between two cases according 
	to the value of \(x + y\).

@@

Setting \(q = S^+(n+1)\) and \(p = S(k)\) in theorem \ref{thm:Stemp} yields the following corollary.

@@

Then, for every \(n \in \mathbb{N}^*\), by using on the last row the coloring \(x \longmapsto g(x - q S(n))\), we have
\[ S(n+k) \geqslant q S(k) + b.\]
\end{proposition}
This proposition corresponds to the fact that the hypotheses made on the coloring of the last row can be weakened because 
the constraints in the definition of a S-template prevent a number from: 
\begin{itemize}
\item being the sum of two of numbers of its own color. 
\item creating a sum in its own color. 
\end{itemize}
The later is not always relevant on the last row given that it is not completely filled.
As a result, we can change the coloring on \([\![ q S(k) + 1,  q S(k) + b]\!]\) (the last row) to extend the previous partitions.
\newline

@@

A S-template with \(n\) colors is said to be symmetric if the partition into \(n\) sum-free subsets derived 
from this template by applying the extension procedure of theorem \ref{thm:Stemp} with a sum-free partition of length 1 is symmetric. 
\end{definition}

We produced S-templates using a SAT solver \cite{Lingeling2017}, hence providing lower bound on \(S^+\) and inequalities 
of the type \(S(n+k) \geqslant a S(n) + b\). We sought templates providing us with the largest value of 
\((a, b)\) (in the lexicographic order). In order to reduce the search space when the number of colors exceeded 
five we 
only looked for symmetric S-templates, we assumed that the special color was the last color to appear and we constrained 
the \(m_c\)'s out of being too small. Details concerning the encoding as a SAT problem can be found in \cite{Heule2017}.

\par
The following six inequalities are given by the current best S-templates with \(n \leqslant 7\) colors.
\begin{align}
	S(n + 1) \geqslant &~3\,S(n)  + 1 \label{S(n+1)}\\
	S(n + 2) \geqslant &~9\,S(n)  + 4 \label{S(n+2)}\\
	S(n + 3) \geqslant &~33\,S(n) + 6 \label{S(n+3)}
\end{align}

@@

\par The \hyperref[S-templates]{templates} corresponding to inequalities (\ref{S(n+3)}), (\ref{S(n+4)}) and 
(\ref{S(n+5)}) are listed in Appendix \ref{S-templates}.

\par
Inequalities (\ref{S(n+1)}), (\ref{S(n+2)}) and (\ref{S(n+3)}) cannot be further improved (with this definition of S-template). 
Inequality (\ref{S(n+4)}) cannot be further improved by only searching for symmetric S-templates whose special color is the 
last in the order of apparition (and with a multiplicative factor less than or equal to 118). Inequality (\ref{S(n+5)}) can most 
likely be further improved but the improvement probably will not be substantial. Finally, a S-template corresponding to inequality 
(\ref{S(n+6)}) was found by extending into a S-template the Schur 6 partition used in \cite{rowley2021improved} (owing to the 
size of this template, its interest is limited and since it can easily be derived from above mentioned partition, it is not given in 
Appendix \ref{S-templates}). Although we could not find a better S-template with seven colors, inequality (\ref{S(n+6)}) can definitely be 
improved by a wide margin.

@@

Except for \(S(8)\), \(S(9)\) and \(S(13)\), the best lower bounds are obtained thanks to
the fifth inequality \( S(n+5) \geqslant 380 \, S(n) + 148\). The table
does not go any further, but the same inequality allows one to improve the
lower bounds for \( n \geqslant 15 \).

@@

First, we formalized Rowley's template-based constructions \cite{RowleyRamsey} in the context of Schur numbers 
by introducing S-templates as well as a new sequence, \(S^+\). We found new S-templates allowing us to obtain 
new lower bounds for schur numbers. One may notice that we mostly gave only lower bounds for \(S^+\). It should be possible to 
find better S-templates by making different assumptions or using a different method (Monte-Carlo methods, for instance).

\par
In the next section, we provide similar results for weak Schur numbers. We introduce WS-templates and a corresponding sequence, 
\(\WS^+\). Then, we derive similar relations and a construction method allowing us to find new lower bounds for weak Schur numbers.
